---
title: "Alzheimer_Project.Rmd"
author: "Ryan S"
date: "2022-08-01"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Scatterplot"
author: "Ryan S"
date: "2022-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loads the necessary packages and variables for the project. Also saves the alzheimer dataset to a simpler name 'alz'
```{r}
library(ggplot2)
library(tidyverse)
library(janitor)
library(dplyr)
library(plot.matrix)
library(expss)
library(nnet)
alzheimer <- read.csv("alzheimer_data.csv")
alzheimer<-mutate(alzheimer, thippo = lhippo + rhippo)
vthippo <- alzheimer$thippo
alzheimer <- mutate(alzheimer, rthippo = thippo/naccicv)
alzheimer$rthippo
rthippo1 <- alzheimer$tiratio
alz <- alzheimer
```


```{r}
ggplot(alzheimer, aes(x = age, y=rthippo, color = diagnosis))+
  geom_point() +
  geom_smooth(method='lm')
```

```{r}

alz$AD <- factor(alz$diagnosis, labels=c('NO', 'YES', 'YES'))

ggplot(alz, aes(x=age, y=rtent, color = AD))+
  geom_point (size = 2, shape = 20,
              alpha = 2)+
  geom_smooth(method=lm)+
  theme_bw()+
    scale_color_brewer(palette = "Dark2",
                     type = qual)+
  ylab("Entorhinal volume over total volume") + ggtitle("Sccaterplot of Rtent in association to Age in relation to diagnosis") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Age Of Patients")
```

```{r}
alzheimer <- mutate(alzheimer, tparcort = lparcort + rparcort)
```

## Creates a bar graph which compares diagnosis with gender (male and female). 
```{r}
ggplot(alzheimer, aes(x=diagnosis, fill = factor(female)))+
  geom_bar(size = 2,
             alpha = 2, position="fill")+
  scale_color_brewer(palette = "Pastel1",
                     type = qual)+ #looks at different colors
  theme_bw()
```

## Creates a scatterplot which compares left hippocampus volume to how old the patient is, and how the older a subject gets,
## the smaller the left hippocampus gets.
```{r}
fem <- factor(alzheimer$female, labels=c('male', 'female'))
ggplot(alzheimer, aes(x=age, y=lhippo, color = fem))+
  geom_point(size = 2, shape = 16,
             alpha = 10)+
  geom_smooth(method=lm)+
  scale_color_brewer(palette = "Accent",
                     type = qual)+
  theme_bw()
```

```{r}
chisq.test(fem, y = alzheimer$diagnosis)
```

```{r}
alzheimer<-mutate(alzheimer, tparcort =lparcort + rparcort)
vtparcort<-alzheimer$tparcort
alzheimer<-mutate(alzheimer, ttempcor =ltempcor + rtempcor)
vttempcor<-alzheimer$ttempcor
alzheimer<-mutate(alzheimer, tcac =lcac + rcac)
vtcac<-alzheimer$tcac
alzheimer<-mutate(alzheimer, tent =lent + rent)
vtent<-alzheimer$tent
alzheimer<-mutate(alzheimer, tparhip =lparhip + rparhip)
vtparhip<-alzheimer$tparhip
alzheimer<-mutate(alzheimer, tposcin =lposcin + rposcin)
vtposcin<-alzheimer$tposcin

ran <- sample(1:2700,30)
new_data <- alzheimer[ran,]
pairs(~thippo + frcort + tparcort + ttempcor + tcac + tent + tparhip + tposcin,
      data=new_data,
      main="Simple Scatterplot Matrix")

```

```{r}
diagonises <- factor(alzheimer$diagnosis, labels=c('No Ad', 'Has AD', 'Has AD'))
ggplot(alzheimer, aes(x=rthippo, y = ..density..,
                      fill = factor(diagonises)))+
  geom_histogram(size = .2,
             alpha = 2)+
  scale_color_brewer(palette = "Pastel1",
                     type = qual)+ #looks at different colors
  theme_bw()
```

```{r}
AD <- factor(alzheimer$diagnosis, labels=c('No', 'Yes', 'Yes'))
AD <- factor(AD)
ggplot(alz, aes(x=rtposcin)) +
  geom_histogram(aes(y=..count../sum(..count..),fill = AD),
                 alpha = .4)+
  scale_color_manual(values = c("red","blue"))+
  theme_bw() + geom_vline(xintercept = mean2, linetype="longdash", 
                color = "red", size=1.5) + geom_vline(xintercept = mean1, linetype="longdash", 
                color = "blue", size=1.5) + xlab("posterior cingulate over total volume")
```

```{r}
tempvar1 = filter(alz, alz$diagnosis != 0)
tempvar0 = filter(alz, alz$diagnosis == 0)
mean1 <- mean(tempvar1$rtposcin)
mean2 <- mean(tempvar0$rtposcin)

t.test(tempvar0$tiratio)
t.test(tempvar0$ttempcor)
t.test(tempvar0$frcort)
t.test(tempvar0$tparcort)
t.test(tempvar0$tcac)
t.test(tempvar0$tent)
t.test(tempvar0$tparhip)
t.test(tempvar0$tposcin)
t.test(tempvar0$thippo)
t.test(tempvar0$tiratio)
t.test(tempvar0$tiratio)

(t.test(alz$rtparhip ~ alz$AD))
(t.test(alz$rtent ~ alz$AD))

t.test(alz$rthippo ~ alz$AD)
t.test(alz$age ~ alz$AD)
t.test(alz$educ ~ alz$AD)
t.test(alz$rfrcort ~ alz$AD)
t.test(alz$rttempcor ~ alz$AD)
t.test(alz$rtcac ~ alz$AD)
t.test(alz$rtposcin ~ alz$AD)
t.test(alz$rparcort ~ alz$AD)


```

```{r}
examplematrix <- matrix(runif(35), ncol=5) # create a numeric matrix object
class(examplematrix)
par(mar=c(5.1, 4.1, 4.1, 4.1)) # adapt margins
plot(examplematrix)

```

```{r}
icvolume <- alzheimer$naccicv

#Ttempcor
rttempcor1<-vttempcor/icvolume
alz <- mutate(alz, rttempcor = rttempcor1)
#Tparcort
rtparcort1<-vtparcort/icvolume
alz <- mutate(alz, rtparcort = rtparcort1)
#FrCort
rfrcort1<-alzheimer$frcort/icvolume
alz <- mutate(alz, rfrcort = rfrcort1)
#Tcac
rtcac1<-vtcac/icvolume
alz <- mutate(alz, rtcac = rtcac1)
#Tent
rtent1<-vtent/icvolume
alz <- mutate(alz, rtent = rtent1)
#TParhip
rtparhip1<-vtparhip/icvolume
alz <- mutate(alz, rtparhip = rtparhip1)
#Tposcin
rtposcin1<-vtposcin/icvolume
alz <- mutate(alz, rtposcin = rtposcin1)


```

```{r}
library(ggplot2)
ggplot(data = alzheimer, aes(x=age, y=female, fill=AD)) + 
  geom_tile()
```
```{r}
alz<-mutate(alz, tparcort =lparcort + rparcort)
vtparcort<-alzheimer$tparcort
alz<-mutate(alz, ttempcor =ltempcor + rtempcor)
vttempcor<-alzheimer$ttempcor
alz<-mutate(alz, tcac =lcac + rcac)
vtcac<-alzheimer$tcac
alz<-mutate(alz, tent =lent + rent)
vtent<-alzheimer$tent
alz<-mutate(alz, tparhip =lparhip + rparhip)
vtparhip<-alzheimer$tparhip
alz<-mutate(alz, tposcin =lposcin + rposcin)
vtposcin<-alzheimer$tposcin
alz<-mutate(alz, thippo = lhippo + rhippo)
vthippo <- alzheimer$thippo
alz <- mutate(alz, tiratio = thippo/naccicv)
tiratio1 <- alzheimer$tiratio
alz_temp <- mutate(alz, ages <- alzheimer$age)
alz_temp <- mutate(alz, education <- alzheimer$educ)
```

```{r}
alz_temp <- data.frame(alz_temp)
alz_temp <- alz_temp %>% select(-starts_with("rhippo"))
```

```{r}
corr <- round(cor(alz_temp), 1)
head(corr[, 1:6])
p.mat <- cor_pmat(alz_temp)
head(p.mat[, 1:4])
ggcorrplot(corr, hc.order = TRUE, outline.col = "white", method = "square", type = "lower", ggtheme = ggplot2::theme_gray, lab = "TRUE") + ggtitle("                   Correlation Matrix")
```

```{r}
#Ttempcor
vhippo<-alz$thippo/icvolume
alz<-mutate(alz, rthippo = vhippo)
vttempcor<-alz$ttempcor/icvolume
alz <- mutate(alz, rttempcor = vttempcor)
vttempcor<-alz$ttempcor/icvolume
alz <- mutate(alz, rttempcor = vttempcor)
#Tparcort
vtparcort<-alz$tparcort/icvolume
alz <- mutate(alz, rparcort = vtparcort)
#FrCort
vfrcort<-alz$frcort/icvolume
alz <- mutate(alz, rfrcort = vfrcort)
#Tcac
vtcac<-alz$tcac/icvolume
alz <- mutate(alz, rtcac = vtcac)
#Tent
vtent<-alz$tent/icvolume
alz <- mutate(alz, rtent = vtent)
#TParhip
vtparhip<-alz$tparhip/icvolume
alz <- mutate(alz, rtparhip = vtparhip)
#Tposcin
vtposcin<-alz$tposcin/icvolume
alz <- mutate(alz, rtposcin = vtposcin)

icvolume <- alz$naccicv
```

```{r}
summary(lm(tiratio ~ age + AD, data = alz))

summary(lm(rthippo ~ age * AD, data = alz))

summary(lm(alz_temp$rthippo ~ age + AD, data = alz_temp))

lm(alz_temp$rthippo ~ alz_temp$age + educ, data = alz_temp)

summary(glm(formula = AD ~ tiratio + age + educ, family = "binomial", data = alz_temp))

chisq.test(alz_temp$tiratio)
```


```{r}
linear_modeling <-function(mydata, response){
summary(lm(response ~ age * AD, data = mydata))
}



linear_modeling(alz_temp, alz_temp$rthippo)
linear_modeling(alz, alz_temp$rttempcor)
linear_modeling(alz, alz_temp$rfrcort)
linear_modeling(alz, alz_temp$rtcac)
linear_modeling(alz, alz_temp$rtent)
linear_modeling(alz, alz_temp$rtparhip)
linear_modeling(alz, alz_temp$rtposcin)
linear_modeling(alz, alz_temp$rparcort)





```

```{r}
data <- alz

data <- data %>%
  mutate(disease_status = ifelse(diagnosis == 0, yes = 0, no = 1)) %>%
  mutate(across(c(disease_status, diagnosis, female), as.factor))

is.factor(data$disease_status) 

table(data$disease_status,data$diagnosis)


train <- data %>% slice_sample(n = 0.7*nrow(data))
test <- data %>% filter(!(id %in% train$id))

# Linear regression model

lin_mod <- glm(lhippo ~ age + educ + rthippo + rfrcort + rtcac + rtent + rtparhip + rtposcin + rparcort + rttempcor, data = train)
test_pred <- predict(lin_mod, newdata = test)
plot(test$lhippo, test_pred)


# Logistic regression model

logistic_mod <- glm(disease_status ~ age + educ + rthippo + rfrcort + rtcac + rtent + rtparhip + rtposcin + rparcort + rttempcor, family = binomial,
                    data = data)

logistic_mod <- glm(disease_status ~ age + educ + rthippo + rfrcort + rtcac + rtent + rtparhip + rtposcin + rparcort + rttempcor, family = binomial,
                    data = train)

test_prob <- predict(logistic_mod, newdata = test, type = 'response')

test <- test %>%
  mutate(predicted_class = ifelse(test_prob > 0.5, 1, 0))


table(test$disease_status, test$predicted_class)

cross_cases(test, disease_status, predicted_class)

mean(test$disease_status == test$predicted_class)

```

```{r}
multinom(forumula = AD, data = alz)
data$diagnosis <- relevel(data$diagnosis, ref="0")
mymodel <- multinom(data$diagnosis~., data = alz_temp)
```


```{r Knit}
data <- alz

data <- data %>%
  mutate(disease_status = ifelse(diagnosis == 0, yes = 0, no = 1)) %>%
  mutate(across(c(disease_status, diagnosis, female), as.factor))

is.factor(data$disease_status) 

table(data$disease_status,data$diagnosis)


train <- data %>% slice_sample(n = 0.7*nrow(data))
test <- data %>% filter(!(id %in% train$id))

# Linear regression model




# Logistic regression model

multinom_mod <- multinom(diagnosis ~ age + educ + rthippo + rfrcort + rtcac + rtent + rtparhip + rtposcin + rparcort + rttempcor,  data = data)

summary(multinom_mod)

multinom_mod <- multinom(diagnosis ~ age + educ + rthippo + rfrcort + rtcac + rtent + rtparhip + rtposcin + rparcort + rttempcor,  data = train)

summary(multinom_mod)

test_pred <- predict(multinom_mod, newdata = test, type = 'class')

table(test$diagnosis, test_pred)

mean(test$diagnosis == test_pred)

```

```{r}
graphed_data <- function(xval, label){
ggplot(alzheimer, aes(x=xval)) + 
    geom_histogram(aes(y=..count../sum(..count..),  fill = factor(diagonises)),alpha = 0.4)+
    scale_color_brewer(palette = "Pastel1", type = qual)+
    theme_bw()+
    geom_vline(xintercept = mean2, color = "red", size=1.5, linetype = "longdash")+
    geom_vline(xintercept = mean1, color = "blue", size=1.5, linetype = 5)+
    xlab(label)+
    ylab("Percent of total data in each bar")
}
```